### 2017-03-03 sections/implementation.tex
Update on Overleaf.

From the properties of SVD method presented in section \ref{sec:math} it follows the fact that a matrix can be represented in the form of its SVD components as a sum of rank 1 matrices in the form

@@

\begin{figure}[ht]
\centering\includegraphics[width=0.7\textwidth]{figures/low_rank_decomposition_diagram}
\caption{Decomposition of input matrix $\mtrx{A}$ into diagonal matrix of singular values $\mtrx{\Sigma}$ and matrices of left and right singular vectors. Light color illustrates low-rank approximation.}
\label{fig:lowrank_svd}
\end{figure}

>>>>

From the definition of SVD in (\ref{eq:svd-def}) and from the properties of SVD presented in section \ref{sec:math} follows the fact that a matrix can be represented in the form of its SVD components as a sum of rank 1 matrices in the form

@@

\begin{figure}[H]
\centering\includegraphics[width=0.7\textwidth]{figures/low_rank_decomposition_diagram}
\caption{Decomposition of input matrix $\mtrx{A}$ into diagonal matrix of singular values $\mtrx{\Sigma}$ and matrices of left and right singular vectors. Light color illustrates low-rank approximation.}
\label{fig:lowrank_svd}
\end{figure}

### 2017-03-03 sections/math.tex
Update on Overleaf.

\todo[inline]{Singular Value Decomposition in general}

% Describe how to calculate decomposition of arbitrary matrix with example

@@

\todo[inline]{Singular values and norms of matrix}

(following taken from \cite{SairaBanu2015})

Mean square error:
\begin{equation}
MSE=\frac{1}{m n} \sum_{i=1}^{m} \sum_{j=1}^{n} (A_{ij} - A'_{ij})^{2}
\label{eq:mse-def}
\end{equation}

where \textit{A} represents the original image and \textit{A'} represents the
reconstructed image of dimension $m \times n$.

Rooted Mean Square Deviation:
\begin{equation}
RMSD=\sqrt{MSE}
\label{eq:rmsd-def}
\end{equation}

Normalized Rooted Mean Square Deviation:
\begin{equation}
NRMSD=\frac{RMSD}{X_{max}-X_{min}}=\frac{\sqrt{MSE}}{X_{max}-X_{min}}
\label{eq:nrmsd-def}
\end{equation}

Peak signal to noise ratio [db]:
\begin{equation}
PSNR=10\log_{10}\frac{(X_{max}-X_{min})^{2}}{MSE}
\end{equation}

@@

\subsection{Sparse matrix of details}

\subsection{Randomized SVD}

>>>>

% TODO: Singular Value Decomposition in general
% TODO: Describe how to calculate decomposition of arbitrary matrix with example
% TODO: add more citations separated by comma

Singular value decomposition (SVD) \cite{Baker2013, Kalman2002} is based on a theorem from linear algebra which says that a rectangular matrix $\mtrx{A}$ can
be broken down into the product of three matrices - an orthogonal matrix $\mtrx{U}$, a diagonal
matrix $\mtrx{S}$, and the transpose of an orthogonal matrix $\mtrx{V}$:

\begin{equation}
\mtrx{A_{m \times n}} = \mtrx{U_{m \times m}} \mtrx{S_{m \times n}} \mtrx{V^{T}_{n \times n}},
\label{eq:svd-def}
\end{equation}

where $\mtrx{U^{T}U} = \mtrx{I}$, $\mtrx{V^{T}V} = \mtrx{I}$; the columns of $\mtrx{U}$ are orthonormal eigenvectors of $\mtrx{AA^{T}}$ , the
columns of $\mtrx{V}$ are orthonormal eigenvectors of $\mtrx{A^{T}A}$, and $\mtrx{S}$ (sometimes referred to as $\mtrx{\Sigma}$) is a diagonal matrix containing singular values in descending order which are at the same time the square roots of eigenvalues from $\mtrx{U}$ or $\mtrx{V}$.

SVD can be seen as a method for transforming correlated variables into a set of uncorrelated ones that better expose the various relationships among the original data items. At the same time, SVD is a method for identifying and ordering the dimensions along which data points exhibit the most variation. Once we have identified where the most variation is, it is possible to find the best approximation of the original data points using fewer dimensions. Hence, SVD can be seen as a method for data reduction/compression.

This is the basic idea behind SVD: taking a high dimensional, highly variable set of data points and reducing it to a lower dimensional space that exposes the substructure of the original data more clearly and orders it from most variation to the least. What makes SVD practical for data compression applications is that you can simply ignore variation below a particular threshold to massively reduce your data but be assured that the main relationships of interest have been preserved.

@@

% TODO: Singular values and norms of matrix

The objective of data compression is to represent an input with lesser amount of data. However, the data reconstructed from its compressed representation may or may not be able to recreate the exact copy of original image. A compression technique can be lossy or lossless based on the quality of data it restores.

Low-rank approximation matrix method that is described in this paper is a lossless compression technique. Several error metrics are used to control the quality of results \cite{SairaBanu2015}.

\begin{itemize}
\item \textbf{Mean square error}
\begin{equation}
MSE=\frac{1}{m n} \sum_{i=1}^{m} \sum_{j=1}^{n} (A_{ij} - A'_{ij})^{2},
\label{eq:mse-def}
\end{equation}

where $\mtrx{A}$ represents the original image and $\mtrx{A'}$ represents the
reconstructed image of dimension $m \times n$.

\item \textbf{Rooted Mean Square Deviation}
\begin{equation}
RMSD=\sqrt{MSE}
\label{eq:rmsd-def}
\end{equation}

\item \textbf{Normalized Rooted Mean Square Deviation}
\begin{equation}
NRMSD=\frac{RMSD}{X_{max}-X_{min}}=\frac{\sqrt{MSE}}{X_{max}-X_{min}},
\label{eq:nrmsd-def}
\end{equation}

where $X_{min}$ and $X_{max}$ are elements of input matrix $\mtrx{A}$ with minimum and maximum value, respectively.

\item \textbf{Peak signal to noise ratio}

PSNR is most commonly used to measure the quality of reconstruction of lossy compression methods (e.g., image compression). The signal in this case is the original data, and the noise is the error introduced by compression. When comparing compression methods, PSNR is an approximation to human perception of reconstruction quality. PSNR is usually expressed in terms of the logarithmic decibel scale (dB).

\begin{equation}
PSNR=10\log_{10}\frac{(X_{max}-X_{min})^{2}}{MSE}
\end{equation}

@@

\end{itemize}

\subsection{Randomized SVD}
In \cite{Holmes, Halko2011, Candes2009} are described randomized algorithms for constructing approximate matrix factorizations. Their main purpose is to accelerate calculations of matrix decompositions which tend to be very slow for large data sets.

Exact SVD of a $m \times n$ matrix has time complexity $O(min(mn^2, m^2n))$ using the "big-O" notation. Assume the data points are in the columns of $\mtrx{A} \in \mtrx{M_{m,n}}(\mathbb{R})$ where $m \leq n$. Note that $\mtrx{AA^T}$ is the dataset covariance matrix. Then a simple method is to randomly choose $k<m$ columns of $\mtrx{A}$ that form a matrix $\mtrx{B}$. Statistically, the SVD of $\mtrx{BB^T}$ will be close to that of $\mtrx{AA^T}$; thus it suffices to calculate the SVD of $\mtrx{B}$, the complexity of which, is only $O(k^2m)$.

% http://mathoverflow.net/questions/161252/what-is-the-time-complexity-of-truncated-svd

% subsection: Sparse matrix of details: Mozna uplne vyhodit, stejne to neni implementovano -- uvest jen v sekci Optimization nebo v Conclusion
